---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "README-"
)
```

# rabi
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rabi)](https://cran.r-project.org/package=rabi) 

The goal of `rabi` is to facilitate the design and generation of color (or symbol) codes that can be used to mark and identify individual animals (specifically aimed for research projects). These codes can be selected such that the IDs are robust to partial erasure: even if parts of the code are lost, the entire identity of the animal can be reconstructed Thus, animal subjects are not confused and no ambiguity is introduced.

## Installation
You can install `rabi` from CRAN with:

```{r cran-installation, eval = FALSE}
install.packages("rabi")

```

You can also install `rabi` from Github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("andburch/rabi")

```

## Example

Although these methods and principles are applicable across taxa and situations, we will demonstrate their usefulness on ants as an illustration. A well-thought-out identification system is important in social insect research because colonies are often composed of hundreds to thousands  of virtually identical workers. Additionally, their tendency to groom themselves (and each other) creates a situation where markings are especially susceptible to becoming detached and disrupting identification.

Let's say we have 100 ants we need to track and observe: we need to *always* know their individual identities. We have five colors of paint which we will apply as drops to locations on their backs. Their heads and thoraxes will each receive a colored drop and then they will get two drops on their abdomens (i.e. gasters). We want a coding scheme that will still perfectly identify the ants even if they _all_ lose one of these marks.


```{r example}

total.length <- 4  #we have four places to put paint,
redundancy <- 1    #we want robustness to single erasures,
alphabet <- 5      #and we currently have five types of paint in stock
 #rs_IDs() is one of the functions that can generate codes
codes <- rabi::rs_IDs(total.length, redundancy, alphabet)
 #prep the output to be displayed in a table
codes <- t(do.call("cbind",codes))
knitr::kable(
  head(codes, n = 10L), 
  col.names = c("Head","Thorax","R. Abdomen","L. Abdomen"), 
  align = "c", caption = "ID sequences for ants")


```


We can easily add our paint color names to the output as well. (However, I would personally save the mapping for future reference, etc.)
```{r more-example, message=FALSE}
paint.names <- c("red","light blue", "greenish", "off-white", "yellow")
color.codes <- rabi::codes_to_colors(codes, paint.names)
 #prep the output to be displayed in a table
color.codes <- t(do.call("cbind",color.codes))
knitr::kable(
  head(color.codes, n = 10L), 
  col.names = c("Head","Thorax","R. Abdomen","L. Abdomen"), 
  align = "c", caption = "ID paint sequences for ants")


```
```{r keep-the-message, echo=FALSE}
color.codes <- rabi::codes_to_colors(codes, paint.names)
```


##More
For a (possibly buggy) Shiny-based GUI suited for new R users, try running this command.
```{r to-GUI, eval = FALSE}
rabi::exampleGUI()
```


For additional functions, deeper examples, and such, check out the vignette on designing color band schemes for bird legs: it's chock full of cool stuff.
```{r to-vignette, eval = FALSE}
utils::vignette('loosebirdtag', package='rabi')
```

Lastly, check out our journal article on this subject:

> Burchill, A. T., & Pavlic, T. P. (2019). Dude, where's my mark? Creating robust animal identification schemes informed by communication theory. *Animal Behaviour*, 154, 203-208. [doi:10.1016/j.anbehav.2019.05.013](https://doi.org/10.1016/j.anbehav.2019.05.013)
 
