---
title: "Tweaking a Coding Scheme: for Bird Tagging"
author: "Andrew Burchill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tweaking a Coding Scheme: for Bird Tagging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this vignette, we will tweak a color coding scheme generated by the `reed_solomon()` function. In our particular hypothetical situation, we plan on putting color bands around the legs of some bird. We have eight different band colors, and we want to put five bands on each bird: two bands below each of the bird's "ankles" (around the tarsometatarsus) and one band above the bird's right "ankle" joint (around the right tibiotarsus). A horrible "graphical" representation can be seen below.

```{r, echo=FALSE, collapse=TRUE,comment=""}
cat("     |   |            |  {5}\n     o   o            o   o\n     |   |     =>    {1} {3}\n     .   .           {2} {4}\n    /|\\ /|\\          /|\\ /|\\")

```

Let us say that these bands fall off somewhat frequently; we want our coding scheme to be robust to the loss of two bands. If these bands occupy unique positions and can't be moved around, we can use the `reed_solomon()` function without further modification.
```{r}
alphabet <- 8      # the number of colors we have
total_length <- 5  # the number of positions we want band
redundancy <- 2    # how many bands we can lose but still ID perfectly

codes <- rabi::reed_solomon(total_length, redundancy, alphabet)
head(codes, n = 4L)

```
Congratualtions! It's as simple as one line of code.

##It's not so simple

However, things get a bit trickier. In reality, these bands may be realtively loose around the legs of the bird. The unique positions of bands `{1}` and `{3}` are probably only maintained due to the physical presence of bands `{2}` and `{4}`. If `{2}` or `{4}` were to somehow be removed, the band above them would succumb to gravity and descend into the newly vacated position. In the below "diagram," we see that a loss of both `{1}` and `{2}` will yield the same result. In this case, it would be impossible to tell which original position the remaining band represented. 
```{r, echo=FALSE, collapse=TRUE,comment=""}
cat("     |  {5}           |  {5}           |  {5}\n     o   o            o   o            o   o\n     |  {3}   AND     |  {3}    =>     |  {3}\n    {1} {4}          {2} {4}          {?} {4}\n    /|\\ /|\\          /|\\ /|\\          /|\\ /|\\")

```
It becomes clear that if a color band is lost

```{r}
codes <- rabi::reed_solomon(total_length = 5, redundancy = 2, alphabet = 8)
 #turn the list of codes into a matrix for ease of editing
codes <- t(do.call("cbind",codes))  
 #now only select the codes where {1} and {3} are not smaller than their lower partners {2} and {4}
codes <- codes[which(codes[,1] >= codes[,2] & codes[,3] >= codes[,4]), ]
 #turn it back into a list...
codes <- split(codes, 1:nrow(codes))
names(codes) <- NULL
head(codes, n = 4L)

```
Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
